---
title: "population pyramids: an animated example"
output: html_notebook
---

Population pyramids are one of the familiar data viz options in the demographer's tool box.

[Kyle Walker](http://personal.tcu.edu/kylewalker/)'s post ["idbr: access the US Census Bureau International Data Base in R"](http://www.arilamstein.com/blog/2016/06/06/idbr-access-us-census-bureau-international-data-base-r) uses international data to create animated population pyramids.

Below, I plagarize the code to create an animated population pyramid for British Columbia.


### The original code

**Notes:** 

* The program requires the `gganimate` package, which is not on [CRAN](https://cran.r-project.org/web/packages/available_packages_by_name.html#available-packages-G) at the time of this writing (2017-05-07). To install the package into your library, follow the directions on the [`gganimate` github page](https://github.com/dgrtwo/gganimate). See also (https://rpubs.com/omicsdata/gganimate).


* You will also need a separate program, Image Magick
https://www.imagemagick.org/script/index.php


Link R to ImageMagick
http://stackoverflow.com/questions/37800390/linking-r-to-imagemagick



* and (although I'm not sure this is essential) the `magick` package 


[CRAN page](https://cran.r-project.org/web/packages/magick/vignettes/intro.html)

[Ropensci entry](https://ropensci.org/blog/2016/08/23/z-magick-release)



#### The packages needed for the entire sequence

```{r}

# data
library(gapminder)
library(idbr) 


# tidyverse
library(tidyverse)
#library(ggplot2)
#library(dplyr)


# animation
library(animation)
library(gganimate)

library(magick)

# THIS IS THE SECRET SAUCE FOR SUCCESSFUL ANIMATION 
# -- embed location of Image Magick executable in R script
ani.options(convert = shQuote('C:/PGMS/ImageMagick-7.0.5-Q16/magick.exe'))

```



a test of a basic gganimate 


```{r}

theme_set(theme_bw())

p <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, color = continent, frame = year)) +
  geom_point() +
  scale_x_log10()

p

gganimate(p)




```


#### A test of the `magick` package

```{r}



# the tiger example
# from https://cran.r-project.org/web/packages/magick/vignettes/intro.html


tiger <- image_read('http://jeroen.github.io/images/tiger.svg')
image_info(tiger)

tiger

# Render svg to png bitmap
image_write(tiger, path = "tiger.png", format = "png")

# convert format
tiger_png <- image_convert(tiger, "png")


```



### Population pyramids

Now we've confirmed that the animation is running, it's time to turn our attention to the creation of a population pyramid in `ggplot2`.


**Note:**  "A Census API key is required.  Obtain one at http://api.census.gov/data/key_signup.html, and then supply the key to the `idb_api_key` function to use it throughout your idbr session."

---


```{r}



# Census API key
idb_api_key("725c5ae0a596294bb619ef881a121c542a89a70d")

male <- idb1('NI', 1990:2050, sex = 'male') %>%
  mutate(POP = POP * -1,
         SEX = 'Male')

female <- idb1('NI', 1990:2050, sex = 'female') %>%
  mutate(SEX = 'Female')

nigeria <- rbind(male, female) 

write_csv(nigeria, "nigeria.csv")


# create the plot
g1 <- ggplot(nigeria, aes(x = AGE, y = POP, fill = SEX, width = 1, frame = time)) +
  coord_fixed() +
  coord_flip() +
  geom_bar(data = subset(nigeria, SEX == "Female"), stat = "identity", position = 'identity') +
  geom_bar(data = subset(nigeria, SEX == "Male"), stat = "identity", position = 'identity') +
  scale_y_continuous(breaks = seq(-5000000, 5000000, 2500000),
                     labels = c('5m', '2.5m', '0', '2.5m', '5m'),
                     limits = c(min(nigeria$POP), max(nigeria$POP))) +
  theme_minimal(base_size = 14, base_family = "Tahoma") +
  scale_fill_manual(values = c('#98df8a', '#2ca02c')) +
  ggtitle('Population structure of Nigeria,') + 
  ylab('Population') +
  xlab('Age') +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  labs(caption = 'Chart by @kyle_e_walker | Data source: US Census Bureau IDB via the idbr R package') + 
  guides(fill = guide_legend(reverse = TRUE))

gg_animate(g1, interval = 0.1, ani.width = 700, ani.height = 600)

# Animate it with gganimate
gganimate(g1, interval = 0.1, ani.width = 700, ani.height = 600)

# save a gif
gganimate(g1, interval = 0.1, ani.width = 700, ani.height = 600, "output.gif")

```


### The British Columbia data

Start with some wrangling

```{r}

BC_pop_1971.2015 <- read_csv("data/BC_Population_1971-2015.csv")

BC_pop_1971.2015

Sex.grp <- c(1, 2)

#BC_pop_1971.2015 <- BC_pop_1971.2015 %>%
#  rename(`90` = `90+`)


BC_pop_1971.2015.sel <- BC_pop_1971.2015 %>%
  select(-c(Total, SGC, `Area Name`, `-1`, Level)) %>%
  filter(Sex %in% Sex.grp)


BC_pop_1971.2015.sel

#  rename(`00` = `0`,
#         `01` = `1`,
#         `02` = `2`,
#         `03` = `3`,
#         `04` = `4`,
#         `05` = `5`,
#         `06` = `6`,
#         `07` = `7`,
#         `08` = `8`,
#         `09` = `9`)


# gather into long format
agegrp <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9",
            "10", "11", "12", "13", "14", "15", "16", "17", "18", "19",
            "20", "21", "22", "23", "24", "25", "26", "27", "28", "29",            
            "30", "31", "32", "33", "34", "35", "36", "37", "38", "39",
            "40", "41", "42", "43", "44", "45", "46", "47", "48", "49",
            "50", "51", "52", "53", "54", "55", "56", "57", "58", "59",
            "60", "61", "62", "63", "64", "65", "66", "67", "68", "69",
            "70", "71", "72", "73", "74", "75", "76", "77", "78", "79",
            "80", "81", "82", "83", "84", "85", "86", "87", "88", "89", "90+")

BC_pop_1971.2015.sel %>%
          gather("0", "1", "2", "3", "4", "5", "6", "7", "8", "9",
                "10", "11", "12", "13", "14", "15", "16", "17", "18", "19",
                "20", "21", "22", "23", "24", "25", "26", "27", "28", "29",            
                "30", "31", "32", "33", "34", "35", "36", "37", "38", "39",
                "40", "41", "42", "43", "44", "45", "46", "47", "48", "49",
                "50", "51", "52", "53", "54", "55", "56", "57", "58", "59",
                "60", "61", "62", "63", "64", "65", "66", "67", "68", "69",
                "70", "71", "72", "73", "74", "75", "76", "77", "78", "79",
                "80", "81", "82", "83", "84", "85", "86", "87", "88", "89", "90+",
                key = "age", value = "population")

BC_pop_1971.2015.long <- BC_pop_1971.2015.sel %>%
          gather(`0`, `1`, `2`, `3`, `4`, `5`, `6`, `7`, `8`, `9`,
                `10`, `11`, `12`, `13`, `14`, `15`, `16`, `17`, `18`, `19`,
                `20`, `21`, `22`, `23`, `24`, `25`, `26`, `27`, `28`, `29`,            
                `30`, `31`, `32`, `33`, `34`, `35`, `36`, `37`, `38`, `39`,
                `40`, `41`, `42`, `43`, `44`, `45`, `46`, `47`, `48`, `49`,
                `50`, `51`, `52`, `53`, `54`, `55`, `56`, `57`, `58`, `59`,
                `60`, `61`, `62`, `63`, `64`, `65`, `66`, `67`, `68`, `69`,
                `70`, `71`, `72`, `73`, `74`, `75`, `76`, `77`, `78`, `79`,
                `80`, `81`, `82`, `83`, `84`, `85`, `86`, `87`, `88`, `89`, `90+`,
                key = "age", value = "population")

BC_pop_1971.2015.long


```


```{r}

g2 <- ggplot(BC_pop_1971.2015.long, aes(x = age, y = population, fill = Sex, width = 1, frame = Year)) +
  coord_fixed() +
  coord_flip() +
  geom_bar(data = subset(BC_pop_1971.2015.long, Sex == "2"), stat = "identity", position = 'identity') +
  geom_bar(data = subset(BC_pop_1971.2015.long, Sex == "1"), stat = "identity", position = 'identity') +
  ggtitle('Population structure of British Columbia,') 



g2 <- ggplot(BC_pop_1971.2015.long, aes(x = age, y = population, fill = Sex, width = 1, frame = Year)) +
  coord_fixed() +
  coord_flip() +
  geom_bar(data = subset(BC_pop_1971.2015.long, Sex == "2"), stat = "identity", position = 'identity') +
  geom_bar(data = subset(BC_pop_1971.2015.long, Sex == "1"), stat = "identity", position = 'identity') +
  scale_y_continuous(breaks = seq(-50000, 50000, 10000),
                     labels = c('50000', '40000', '30000', '20000', '10000', '0', 
                                '10000', '20000', '30000', '40000', '50000'),  
                     limits = c(min(-50000), max(50000))) 

+
#  theme_minimal(base_size = 14, base_family = "Tahoma") +
#  scale_fill_manual(values = c('#98df8a', '#2ca02c')) +
  ggtitle('Population structure of British Columbia,') 
# + 
#  ylab('Population') +
#  xlab('Age') +
#  theme(legend.position = "bottom", legend.title = element_blank()) +
#  labs(caption = 'Chart by @kyle_e_walker | Data source: US Census Bureau IDB via the idbr R package') + 
#  guides(fill = guide_legend(reverse = TRUE))


# Animate it with gganimate
gganimate(g2, interval = 0.1, ani.width = 700, ani.height = 600)

# save a gif
gganimate(g2, interval = 0.1, ani.width = 700, ani.height = 600, "output.gif")


```


