---
title: "R Notebook"
output: html_notebook
---

## Slopegraph example

A blog post by Jens Van Bergen at Mountain Doodles, ["Millenials Redux"](http://doodles.mountainmath.ca/blog/2017/08/06/millennials-redux/) got me thinking about the way he represented the change in proprotions...might a slope graph be a better solution?


plagarized from Kyle Walker
https://rpubs.com/walkerke/slopegraph


#### packages

Nothing but the [`tidyverse`](http://www.tidyverse.org/)

* except `readxl` isn't a `tidyverse` auto-load

```{r}

library(tidyverse)
#library(readr)
#library(dplyr)
#library(ggplot2)
#library(magrittr)
library(readxl)

```


---

```{r}

#dat <- read_csv("http://esa.un.org/wpp/ASCII-Data/ASCII_FILES/WPP2012_DB02_POPULATIONS_ANNUAL.csv")
# https://esa.un.org/unpd/wpp/Download/Standard/Population/

#dat <- read_excel("https://esa.un.org/unpd/wpp/DVD/Files/1_Indicators%20(Standard)/EXCEL_FILES/1_Population/WPP2017_POP_F15_1_ANNUAL_POPULATION_BY_AGE_BOTH_SEXES.xlsx")

#dat <- read_excel("../data/WPP2017_POP_F15_1_ANNUAL_POPULATION_BY_AGE_BOTH_SEXES.xlsx", range = cell_rows(c(17, NA)))


#  ** problem: ** replace the "…" with NA


# this is slow...but it works eventually
WPP2017_data <- read_excel("../data/WPP2017_POP_F15_1_ANNUAL_POPULATION_BY_AGE_BOTH_SEXES.xlsx", sheet = "ESTIMATES",
                  range = cell_rows(c(17, NA)), 
                  col_types = c("text", "text", "text", "text",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric",
                                "numeric", "numeric", "numeric", "numeric"
                                )
                  )


saveRDS(WPP2017_data, "../data/WPP2017_data.rds")


ls.str(WPP2017_data)

```


### Additional wrangling 

Necessary to make 2017 file work with the previous code.


#### Already saved the data file? Start here...

```{r}

WPP2017_data <- readRDS("../data/WPP2017_data.rds")

dat <- WPP2017_data

```


#### Wrangling


```{r}
# rename and drop variables
dat1 <- dat %>%
  rename('Location' = 'Region, subregion, country or area *',
         'Time' = 'Reference date (as of 1 July)') 

var_list <- c('0-4','5-9', '10-14', '15-19', '20-24', '25-29',
              '30-34', '35-39', '40-44', '45-49',  
              '50-54', '55-59', '60-64', '65-69', '70-74', '75-79', 
              '80-84', '80+', '85-89', '90-94', '95-99', '100+')

dat1 <- dat1 %>%
  select(-(Index:Variant),
         -(Notes:`Country code`))

ls.str(dat1)

```


* is the problem that "…" is a UTF-8 character (U+2026)?



```{r}

# gather dat1 into long format
dat1_long <- dat1 %>%
  gather(`0-4`:`100+`, key = 'Age_grp', value = 'Pop', na.rm=TRUE)


dat2_long <- dat1_long %>%
  group_by(Location, Time) %>%
  summarise(PopTotal = sum(Pop))


unique(dat2_long$Location)

min(dat2_long$Time)
max(dat2_long$Time)

```





```{r}

# These are the region names in the dataset
regions <- c("AFRICA", "LATIN AMERICA AND THE CARIBBEAN", "NORTHERN AMERICA", "EUROPE", "OCEANIA", "ASIA")




# Filter down to 2015 and 2099 and the medium projection variant, and calculate millions & billions 
dat2 <- dat2_long %>%
#  filter(VarID == 2, Location %in% regions, Time %in% c(2015, 2099)) %>%
  filter(Location %in% regions, Time %in% c(1950, 2015)) %>%
  mutate(millions = round((PopTotal / 1000), 0), billions = round((PopTotal / 1000000), 1)) %>%
  select(Location, Time, millions, billions)


# Shorten a couple region names which I'll be using for labels
dat2$Location <- gsub("AFRICA", "Africa", dat2$Location)
dat2$Location <- gsub("LATIN AMERICA AND THE CARIBBEAN", "Latin America", dat2$Location)
dat2$Location <- gsub("NORTHERN AMERICA", "North America", dat2$Location)
dat2$Location <- gsub("EUROPE", "Europe", dat2$Location)
dat2$Location <- gsub("OCEANIA", "Oceania", dat2$Location)
dat2$Location <- gsub("ASIA", "Asia", dat2$Location)


# Prepare some label columns for the slopegraph
dat2 <- dat2 %>%
  mutate(label15 = ifelse(billions > 1, 
                          paste0(Location, " ", as.character(billions), "b  "),
                          paste0(Location, " ", as.character(millions), "m  ")), 
         label99 = ifelse(billions > 1, 
                          paste0("  ", as.character(billions), "b"), 
                          paste0("  ", as.character(millions), "m")))

dat2

```


```{r}

locs_to_adjust1 <- c("Africa", "Europe", "North America")

locs_to_adjust2 <- c("Asia", "Latin America")

ggplot(dat2) + 
  geom_line(aes(x = as.factor(Time), y = millions, group = Location, color = Location), size = 2) + 
  geom_point(aes(x = as.factor(Time), y = millions, color = Location), size = 5) + 
  theme_minimal(base_size = 18) + 
  scale_color_brewer(palette = "Dark2") + 
  xlab("") + 
  geom_text(data = subset(dat2, Time == 1950 & Location != "Latin America"), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label15), 
            size = 6, hjust = 1) + 
  geom_text(data = subset(dat2, Time == 1950 & Location == "Latin America"), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label15), 
            size = 6, hjust = 1, vjust = 0.8) + 
  geom_text(data = subset(dat2, Time == 2015 & Location %in% locs_to_adjust1), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label99), 
            size = 6, hjust = 0, vjust = 0.8) + 
  geom_text(data = subset(dat2, Time == 2015 & Location %in% locs_to_adjust2), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label99), 
            size = 6, hjust = 0, vjust = 0.2) + 
  geom_text(data = subset(dat2, Time == 2015 & Location == "Oceania"), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label99), 
            size = 6, hjust = 0) + 
  scale_y_log10() + 
  theme(legend.position = "none", 
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(), 
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(), 
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        plot.title = element_text(size = 18)) + 
  ggtitle("Projected population change by region, 1950-2015")


```



---

#### Add the data

Currently approximations based on eyeballing the chart

```{r}

MetroVan_millenials <- read.csv("../data/MetroVan_millenials.csv")

```



And the plot

```{r}

ggplot(MetroVan_millenials) + 
  geom_line(aes(x = as.factor(Year), y = 'pct_25-39', group = Municipality, color = Municipality), size = 2) + 
  geom_point(aes(x = as.factor(Year), y = 'pct_25-39', color = Municipality), size = 5) + 
  theme_minimal(base_size = 18) + 
  scale_color_brewer(palette = "Dark2") + 
  xlab("") + 
  geom_text(data = subset(dat2, Time == 2015 & Location != "Latin America"), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label15), 
            size = 6, hjust = 1) + 
  geom_text(data = subset(dat2, Time == 2015 & Location == "Latin America"), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label15), 
            size = 6, hjust = 1, vjust = 0.8) + 
  geom_text(data = subset(dat2, Time == 2099 & Location %in% locs_to_adjust1), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label99), 
            size = 6, hjust = 0, vjust = 0.8) + 
  geom_text(data = subset(dat2, Time == 2099 & Location %in% locs_to_adjust2), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label99), 
            size = 6, hjust = 0, vjust = 0.2) + 
  geom_text(data = subset(dat2, Time == 2099 & Location == "Oceania"), 
            aes(x = as.factor(Time), y = millions, color = Location, label = label99), 
            size = 6, hjust = 0) + 
  scale_y_log10() + 
  theme(legend.position = "none", 
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(), 
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank(), 
        axis.title.y = element_blank(), 
        axis.text.y = element_blank(), 
        plot.title = element_text(size = 18)) + 
  ggtitle("Projected population change by region, 2015-2099")


```

